Online Balanced Graph Partitioning: Optimizations for Large Clusters

### 2020-05-05 paper/disc20clustering.tex
Intro to LB
wip

In this section we consider a variant of Online Balanced Partitioning, where communication requests are permanent: once a pair of nodes start communicating, they continue forever.
To avoid infinite cost, each communicating pair must be collocated.
We study the instances with finite cost: we assume that the communication graph can be perfectly partitioned in the clusters.
The algorithm's goal is to \emph{learn} the communication graph's partition.

For this setting, we show a surprisingly high lower bound of $\Omega(k \cdot \ell)$.
The lower bound can be easily translated to the partitioning model (Section~\ref{sec:part}) with finite request sequence.
At the end of this section, we discuss an asymptotically optimal upper bound.

@@

In this section, we provide a lower bound $\Omega(k\cdot \ell)$ for the competitive ratio of any deterministic online algorithm.
The construction is oblivious to the choice of the reconfiguration cost $\alpha$: we compare the number of node exchanges.
The lower bound requires $k\geq 3$.
For $k=2$, a lower bound of $3$ is known \cite{repartition-disc}, and we provide a $6$-competitive algorithm (see Section~\ref{sec:k2}).

@@

The adversary constructs ground sets in reaction to choices of a deterministic online algorithm.
We emphasize that the algorithm does not know the ground set before they are revealed.
Once we construct a ground set, it lasts until the end of the input sequence.

The constructed ground sets can be partitioned into our clusters (without splitting nodes).
This way, the optimal algorithm moves to such a partition, and serves all requests for free.
If the algorithm splits any pair of nodes of a ground set, the adversary issue requests to any pair of split nodes.
These two observations imply that the algorithm must eventually collocate nodes of a ground set to be competitive.

@@

We conclude by emphasizing the generality of the lower bound.

\noindent
\textbf{General partitioning problem.}
Every lower bound for the learning problem constitutes a valid lower bound for the partitioning problem.
The bound for partitioning converges to the bound for learning as the input sequence grows.

Our construction allows for an exact lower bound for partitioning with finite input sequence.
To this end, we issue only enough requests until the algorithm collocates the nodes of each ground set.
If it never does, we repeat until it pays $k\cdot \ell$.

>>>>

In this section we consider a variant of Online Balanced Partitioning, where communication requests are permanent: once a pair of nodes start communicating, they continue forever.
To avoid infinite cost, each communicating pair must be collocated.
We study instances with finite cost: we assume that the communication graph can be perfectly partitioned in the clusters.
The algorithm's goal is to \emph{learn} the communication graph, and partition it without too many reconfigurations.

For this setting, we show a surprisingly high lower bound of $\Omega(k \cdot \ell)$.
The lower bound can be easily translated to the partitioning model (studied in Section~\ref{sec:part}) with finite request sequence.
At the end of this section, we discuss an asymptotically optimal upper bound.

@@

We provide a lower bound $\Omega(k\cdot \ell)$ for the competitive ratio of any deterministic online algorithm for the learning problem.
Later we elaborate on how to efficiently transform it to a lower bound for the partitioning problem.
The lower bound requires $k\geq 3$.
For $k=2$, the learning problem is trivial: immediate collocation of a communicating pair is $1$-competitive.
(A partitioning problem for $k=2$ is non-trivial, a lower bound of $3$ is known \cite{repartition-disc}, and we provide a $6$-competitive algorithm (see Section~\ref{sec:k2}).)

@@

The adversary constructs ground sets (a set of nodes that may communicate, see Section~\ref{sec:prelim}) depending on choices of a deterministic online algorithm.
We emphasize that the algorithm does not know ground sets before they are revealed.
Once we construct a ground set, it lasts until the end of the input sequence.

The constructed ground sets can be perfectly partitioned into the clusters (i.e., nodes of each ground set collocated).
This way, the optimal algorithm moves to a perfect partition at the beginning (where requests incur the cost $0$), thus its cost is bounded.
If the algorithm splits any pair of nodes of a ground set, the adversary issue requests to any pair of split nodes.
These two observations imply that the algorithm must eventually collocate all nodes of a ground set to be competitive.

@@

\noindent
\textbf{General partitioning problem.}
Every lower bound for the learning problem constitutes a valid lower bound for the partitioning problem.
The bound for partitioning converges to the bound for learning as the input sequence grows.
However, our construction allows for an exact lower bound for partitioning with finite input sequence.
To this end, we issue only enough requests until the algorithm collocates the nodes of each ground set (if it never does, we terminate after it pays $k\cdot \ell$).
The construction is oblivious to the choice of the reconfiguration cost $\alpha$: we compare the 	number of node exchanges of \ALG and \OPT.
